### 2011-08-09 docs/Bibliografia/bib.tex

\begin{thebibliography}{200}

%Application Notes
\bibitem{ACD} Antenna Circuit Design for RFID Applications, AN710.
\bibitem{ADN} HF Antenna Design Notes, Literature Number: 11-08-26-003.
\bibitem{HFACB} HF Antenna Cook Book, Lit. Number 11-08-26-001.
\bibitem{MRICF} Micore Reader IC Family; Directly Matched Antenna Design, Rev. 2.05, 10 May 2006.
\bibitem{RFIDPA} Mifare®(14443A) 13,56 MHz RFID Proximity Antennas, Application Note.
\bibitem{} MIFARE Type Identification Procedure; AN10833 Rev. 3.1 — 07 July 2009.
\bibitem{} MIFARE ISO/IEC 14443 PICC Selection; AN10834 Rev. 3.0 — 26 June 2009.
\bibitem{} MIFARE Classic Commands; AN10856 Rev. 1.0 — 31 July 2009.
\bibitem{} MIFARE and handling of UIDs; AN10927 Rev. 2.0 — 01 September 2010.
\bibitem{} Mainstream contactless smart card IC for fast and easy; MF1S5009 Rev. 3 — 27 July 2010.
\bibitem{} MIFARE Standard Card IC MF1 IC S50 Functional Specification; Revision 5.1 May 2001.

%Hojas de datos
\bibitem{RC632} MIFARE® and I Code CL RC632 Multiple protocol contactless reader IC.
\bibitem{HD_VLT} 8-BIT BIDIRECTIONAL VOLTAGE-LEVEL TRANSLATOR.
\bibitem{dpy} HD44780U (LCD-II), Dot Matrix Liquid Crystal Display Controller/Driver, Rev. 0.0 1998. 
\bibitem{LDO} LM1117/LM1117I 800mA Low-Dropout Linear Regulator.
\bibitem{LCM} LCM-S01602DSF/A - hoja de datos del LCD16x2.
\bibitem{manualBb} BeagleBoard System Reference Manual Rev C4, Revision 0.0.

%HandBooks
\bibitem{SCHb} Smart Card Handbook, Third Edition, Wolfgang Rankl and Wolfgang Effing, 2003, ISBN: 0-470-85668-8. 
\bibitem{RFIDHb} RFID Handbook: Fundamentals and Applications in Contactless Smart Cards and Identification,
Second Edition, Klaus Finkenzeller, 2003, ISBN: 0-470-84402-7.

%CT/API cap-6
\bibitem{ctapi} Application Independent CardTerminal Application Programming Interface for ICC Applications; 
Deutsche Telekom AG / PZ Telesec, GMD - Forschungszentrum Informationstechnik GmbH, TÜV Informationstechnik GmbH, TELETRUST Deutschland e.V., 30.10.1996.

%CT/BCS cap-6
\bibitem{ctbcs} Application independent CardTerminal Basic Command Set for ICC applications; 
TeleTrusT Deutschland e.V., GMD Forschungszentrum Informationstechnik GmbH, Version 0.9, 28.07.1995.

%IFDhandler
\bibitem{ifdhandler} IFDHandler 3.0 MUSCLE PC/SC IFD Driver API; David Corcoran \& Ludovic Rousseau, July 28, 2004.

%Libros
\bibitem{KiCad} Diseño y desarrollo de circuitos impresos con KiCad, Miguel Pareja Aparicio, 2010, ISBN: 978-84-937769-1-6.

%Normas ISO
\bibitem{7816} Norma ISO/IEC 7816.
\bibitem{14443} Norma ISO/IEC 14443.

%LINKS

\bibitem{OpenPCD} OpenPCD, http://www.openpcd.org/
\bibitem{usb4all} USB4ALL, http://www.fing.edu.uy/inco/grupos/mina/pGrado/pgusb/main.php
\bibitem{librfid} librfid, http://openmrtd.org/projects/librfid/
\bibitem{9G20} GES9G20, http://www.glomationinc.com/product\_9G20.html
\bibitem{Hawk} Hawkboard, http://www.hawkboard.org/
\bibitem{Beagle} Beagleboard, http://beagleboard.org/
\bibitem{KiCad} KiCad, http://www.lis.inpg.fr/realise\_au\_lis/kicad/
\bibitem{gEDA} gEDA, http://www.gpleda.org/index.html
\bibitem{Proxmark} Proxmark, http://www.proxmark.org/proxmark
\bibitem{Angs} Angström, http://www.angstrom-distribution.org/
\bibitem{Angs_MLO} MLO, http://www.angstrom-distribution.org/demo/beagleboard/MLO/
\bibitem{Narc} Narcissus, http://narcissus.angstrom-distribution.org/
\bibitem{termios} termios, http://pubs.opengroup.org/onlinepubs/007908799/xsh/termios.h.html
\bibitem{Make} Make, http://es.wikipedia.org/wiki/Make

%links sobre smart cards y comandos linux necesarios para entender código
\bibitem{} http://ladyada.net/make/simreader/
\bibitem{} http://www.freeinfosociety.com/electronics/schemview.php?id=1995
\bibitem{} http://www.gooze.eu/howto/smartcard-quickstarter-guide
\bibitem{} http://linux.die.net/man/3/dlsym
\bibitem{} http://www.mkssoftware.com/docs/man3/select.3.asp
\bibitem{} http://tldp.org/HOWTO/Program-Library-HOWTO/dl-libraries.html

%links sobre puerto serial y comandos linux necesarios para entender código
\bibitem{} http://www.easysw.com/~mike/serial/serial.html
\bibitem{} http://linux.die.net/man/3/fd\_set
\bibitem{} http://linux.die.net/man/2/select\_tut
\bibitem{} http://www.iearobotics.com/wiki/index.php?title=Tutorial:Puerto\_serie\_en\_Linux
\bibitem{} http://es.tldp.org/COMO-INSFLUG/COMOs/Programacion-Serie-Como/Programacion-Serie-Como-3.html
\bibitem{} http://www.rastersoft.com/articulo/pserie.html
\bibitem{} http://www.faqs.org/docs/Linux-mini/IO-Port-Programming.html
\bibitem{} http://tldp.org/HOWTO/IO-Port-Programming-2.html

%links sobre pcsc
\bibitem{pcsclite} PCSC-Lite, http://pcsclite.alioth.debian.org/
\bibitem{pcsclite_esp} PCSC Workgroup, http://www.pcscworkgroup.com/
\bibitem{pcsc/ccid_down} Descargas PCSC-Lite, https://alioth.debian.org/frs/?group\_id=30105
\bibitem{pcsctools_down} http://ludovic.rousseau.free.fr/softwares/pcsc-tools/
\bibitem{} http://ludovicrousseau.blogspot.com/2011/04/libccid-and-usb-selective-suspend.html
\bibitem{} http://readthedocs.org/docs/pvdevtools\_doc/en/latest/index.html
\bibitem{} http://www.springcard.com/support/pcsclite.html

\end{thebibliography}